### 2023-06-08 Writing/Online-Appendix_supp.rmd

---
title: "Appendix for Paper"
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: true
    reference_docx: "word_template.docx"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ": "
bibliography: PA.bib
csl: field-crops-research.csl
abstract: ""
header-includes:
   - \usepackage{subfig}
---

```{r echo = F, cache = F}
suppressMessages(library(knitr))
suppressMessages(library(here))
suppressMessages(library(officedown))
suppressMessages(library(officer))

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = F,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(lfe)
library(modelsummary)
library(patchwork)
library(ggplot2)
```

```{r figure_setup, cache = F}
theme_update(
  axis.title.x =
    element_text(
      size = 12, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
    ),
  axis.title.y =
    element_text(
      size = 12, angle = 90, hjust = .5, vjust = .9, family = "Times"
    ),
  axis.text.x =
    element_text(
      size = 10, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
    ),
  axis.text.y =
    element_text(
      size = 10, angle = 0, hjust = 1, vjust = 0, family = "Times"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, family = "Times"
    ),
  legend.title =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, family = "Times"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 10),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 12),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r include = FALSE}
#| include: false

# /*===========================================================
#' # Prepare data and results for writing
# /*===========================================================

# "gaussian", "bisquare", "exponential", "tricube", "boxcar"

# kernel_choice <- "gaussian"
# 
# #* read the results rds file
# results <-
#   here("Shared", "Results", kernel_choice, "pi_data.rds") %>%
#   readRDS() %>%
#   .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
#   .[, bias := pi_diff_est - pi_diff] %>%
#   # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
#   .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
#   # === label price ratio
#   .[, pLabelName := "Price Ratio (N/corn)"] %>%
#   .[pRatio == 5.44, pLabel := "Low"] %>%
#   .[pRatio == 6.56, pLabel := "Middle"] %>%
#   .[pRatio == 7.67, pLabel := "High"] %>%
#   .[, pLabel := factor(pLabel, levels = c("Low", "Middle", "High"))]
# 
# #* create figures
# source(here("GitControlled/Codes/3_make_figures_publication.R"))
# assign("g_value_gaussian", g_value)
# assign("g_bias_gaussian", g_bias)
# 
# 
# 
# kernel_choice <- "exponential"
# 
# #* read the results rds file
# results <-
#   here("Shared", "Results", kernel_choice, "pi_data.rds") %>%
#   readRDS() %>%
#   .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
#   .[, bias := pi_diff_est - pi_diff] %>%
#   # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
#   .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
#   # === label price ratio
#   .[, pLabelName := "Price Ratio (N/corn)"] %>%
#   .[pRatio == 5.44, pLabel := "Low"] %>%
#   .[pRatio == 6.56, pLabel := "Middle"] %>%
#   .[pRatio == 7.67, pLabel := "High"] %>%
#   .[, pLabel := factor(pLabel, levels = c("Low", "Middle", "High"))]
# 
# #* create figures
# source(here("GitControlled/Codes/3_make_figures_publication.R"))
# assign("g_value_exponential", g_value)
# assign("g_bias_exponential", g_bias)
```

`r run_pagebreak()`

Economic Evaluation of Geographically Weighted Regression Analysis for Site-specific Nitrogen Management

# Results of Alternative Kernel Functions

There are many different kernel functions to define the distance-decay weightings for GWR model. In the main text, the Gaussian kernel function was employed, which can be mathematically represented as follows:

$$w_{ab}=exp⁡(-0.5[\frac{d_{ab}}{h}]^2 ),$$
where $d_{ab}$ denotes the Euclidean distance between observation $b$ and regression location $a$, $w_{ab}$ denotes the kernel weight assigned to observation $b$, and $h$ denotes the bandwidth.

This appendix compares several other alternative kernel functions, including the Exponential, Boxcar, Bi-square, and Tri-cube, in additional to the Gaussian. The functional forms and graphical representations of these kernels can be found in Figure A1.

The estimation results obtained using alternative kernel functions, apart from the Gaussian, are presented in Figures A2 and A3. For the case of Exponential kernel, which is a continuous kernel like the Gaussian, the optimal bandwidth selection follows a similar corrected AIC approach and results in a similar bandwidth of 18 meters. For the discontinuous kernels of Boxcar, Bi-square, and Tri-cube, however, the 18-meter bandwidth does not encompass a sufficient number of neighboring observations for reliable local regressions, resulting in unrealistic estimation outcomes. To address this issue, an empirical bandwidth of 100 meters is chosen for these discontinuous kernels to enable the inclusion of an adequate number of local observations. Other bandwidth values (such as 80 meters, 150 meters) are also conducted and the resulting estimation outcomes are found to be comparable.

## Comparing the values of GWR-based VRA and SCAM-based URA

Figure A2 shows the distribution of the true economic value of GWR, denoted as the profit difference between the GWR-based VRA (Variable Rate Application) and SCAM-based URA (Uniform Rate Application),
$$\Delta^{GWR}=\Pi^{GWR}-\Pi^{SCAM},$$
which aligns with Figure 3 in the main text but uses alternative kernel functions. The findings observed in Figure 3 under Gaussian kernel, where GWRR performs better as the price ratio increases, and GWRT performs better as the price ratio decreases, hold true across all other kernel function choices. Although there are slight variations in the magnitudes of computed $\Delta^{GWR}$ among different kernel function choices, these differences do not significantly affect the conclusions drawn in the main text.

## Bias in the estimation of the value of GWR-based VRA

Figure A3 shows the distribution of the bias (over-estimation) in the estimated economic value of GWR, measured as the difference between the estimated and true values of GWR-based VRA,
$$bias=\hat{\Delta}^{GWR}-\Delta^{GWR},$$
which aligns with Figure 6 in the main text but uses alternative kernel functions. Similar to the findings of Figure 6 under the Gaussian kernel, the GWRR consistently exhibits significant over-estimation of the value of GWR-based VRA across all other kernel function choices. Also, the GWRT demonstrates smaller over-estimation at low price ratios compared to GWRR, while the over-estimation increases as price ratio rises. The presence of a similar fat tail in the GWRT bias distribution is also evident. The magnitudes of the estimation bias vary across kernel function choices, but the differences are not substantial and do not significantly affect the conclusions drawn in the main text.

Overall, the economic evaluation results of GWR models presented in this study are not significantly influenced by the selection of the kernel function in GWR estimation. Despite slight variations in calculated profit and bias among the different kernels, the qualitative consistency of the final conclusions remains intact regardless of the chosen kernel.

`r run_pagebreak()`

# Figures

```{r, fig.id = "kernel_des", fig.cap = "Figure A1. Different kernel functions for GWR modeling. In the curve diagrams of the right panel, the y-axis is the kernel weight, and the x-axis is the ratio of (d_ab/h)."}
#
```

```{r, fig.id = "pi-dif-dist", fig.cap = "The value of GWR-based VRA over SCAM-based URA for GWRR and GWRT", fig.subcap = c("Gaussian", "Exponential"), fig.height = 5, out.width="50%", fig.asp=1, fig.ncol = 2}
#g_value_gaussian
#g_value_exponential
```

`r run_pagebreak()`

# References

<br />

<div id="refs"></div>

`r run_pagebreak()`

# Appendix