### 2023-06-16 Writing/supplementary_appendix.rmd
finalize appendix; clear the folder;

---
title: "Supplementary Appendix"
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: false
    reference_docx: "word_template_no_section_numbers.docx"
    base_format: "bookdown::word_document2"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure A"
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table A"
       sep: ": "
bibliography: PA.bib
csl: field-crops-research.csl
---

```{r echo = F, cache = F}
suppressMessages(library(knitr))
suppressMessages(library(here))
suppressMessages(library(officedown))
suppressMessages(library(officer))

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = F,
  echo = F,
  fig.cap = TRUE
)
```

```{r}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(lfe)
library(modelsummary)
library(patchwork)
library(ggplot2)
```

```{r figure_setup}
theme_update(
  axis.title.x =
    element_text(
      size = 12, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
    ),
  axis.title.y =
    element_text(
      size = 12, angle = 90, hjust = .5, vjust = .9, family = "Times"
    ),
  axis.text.x =
    element_text(
      size = 10, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
    ),
  axis.text.y =
    element_text(
      size = 10, angle = 0, hjust = 1, vjust = 0, family = "Times"
    ),
  axis.ticks =
    element_line(
      size = 0.3, linetype = "solid"
    ),
  axis.ticks.length = unit(.15, "cm"),
  #--- legend ---#
  legend.text =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, family = "Times"
    ),
  legend.title =
    element_text(
      size = 10, angle = 0, hjust = 0, vjust = 0, family = "Times"
    ),
  legend.key.size = unit(0.5, "cm"),
  #--- strip (for faceting) ---#
  strip.text = element_text(size = 10),
  #--- plot title ---#
  plot.title = element_text(family = "Times", face = "bold", size = 12),
  #--- margin ---#
  # plot.margin = margin(0, 0, 0, 0, "cm"),
  #--- panel ---#
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(fill = NA)
)
```

```{r include = FALSE}
kernel_ls <- c("gaussian", "exponential", "boxcar", "bisquare", "tricube")

gg_appendix <-
  data.frame(kernel = kernel_ls) %>%
  rowwise() %>%
  mutate(results = list(
    here("Shared", "Results", kernel, "pi_data.rds") %>%
      readRDS() %>%
      .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
      .[, bias := pi_diff_est - pi_diff] %>%
      # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
      .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
      # === label price ratio
      .[, pLabelName := "Price Ratio (N/corn)"] %>%
      .[pRatio == 5.44, pLabel := "Low"] %>%
      .[pRatio == 6.56, pLabel := "Middle"] %>%
      .[pRatio == 7.67, pLabel := "High"] %>%
      .[, pLabel := factor(pLabel, levels = c("Low", "Middle", "High"))]
  )) %>%
  #* ===========================================================
  #* Plots of the value of GWR
  #* ===========================================================
  mutate(median_data_value = list(
    results %>%
      .[, .(pi_diff = median(pi_diff)),
        by = c("field_col", "pLabel", "type")
      ]
  )) %>%
  mutate(g_value = list(
    results %>%
      ggplot(data = .) +
      geom_histogram(
        aes(x = pi_diff),
        fill = NA,
        color = "blue",
        bins = 50,
        size = 0.2
      ) +
      geom_vline(
        data = median_data_value,
        aes(xintercept = pi_diff),
        color = "red",
        size = 0.5
      ) +
      geom_text(
        data = median_data_value, color = "red",
        aes(
          x = ifelse(pi_diff < 65, pi_diff + 5, pi_diff - 20),
          y = 150,
          label = paste0("Median = ", round(pi_diff, 2))
        ),
        angle = 0, hjust = -0.1, vjust = 0, size = 3
      ) +
      facet_grid(pLabel ~ type) +
      scale_y_continuous(expand = c(0, 0), limits = c(NA, 200)) +
      scale_x_continuous(breaks = 20 * (-2:3), limits = c(-45, 60)) +
      xlab("The value of VRA over URA ($ per ha)") +
      ylab("Number of Simulation Cases")
  )) %>%
  #* ===========================================================
  #* Plots of the bias in the value of GWR
  #* ===========================================================
  mutate(median_bias_data = list(
    results %>%
      .[, bias := pi_diff_est - pi_diff] %>%
      .[, .(bias = median(bias)),
        by = c("field_col", "pLabel", "type")
      ]
  )) %>%
  mutate(g_bias = list(
    results %>%
      .[, bias := pi_diff_est - pi_diff] %>%
      .[bias < 150, ] %>%
      ggplot(data = .) +
      geom_histogram(
        aes(x = bias),
        fill = NA,
        color = "blue",
        size = 0.2,
        bins = 50
      ) +
      geom_vline(
        data = median_bias_data, aes(xintercept = bias),
        color = "red",
        size = 0.5
      ) +
      geom_text(
        data = median_bias_data, color = "red",
        aes(
          x = ifelse(bias < 65, bias + 2, bias - 85), y = 100,
          label = paste0("Median = ", round(bias, 2))
        ),
        angle = 0, hjust = -0.1, vjust = 0, size = 3
      ) +
      facet_grid(pLabel ~ type) +
      scale_y_continuous(expand = c(0, 0)) +
      xlab("Bias in the Estimation of the Value of VRA over URA ($ per ha)") +
      ylab("Number of Simulation Cases")
  )) %>%
  select(kernel, g_value, g_bias)
```

# Appendix I: Historical Corn and Nitrogen Prices {-}

Figure A\@ref(fig:price) illustrates the historical prices of corn ($p$) and nitrogen fertilizer ($w$), along with the corresponding nitrogen-corn price ratios, during the two-decade period from 2002 to 2022.

```{r, fig.id = "price", fig.cap = "Historical monthly variations in corn price, nitrogen fertilizer price, and the corresponding nitrogen-corn price ratio, 2002-2022"}
knitr::include_graphics("price_ratio_history.png")
```

Corn prices were sourced from the USDA National Agricultural Statistics Service [@NASS2022] in monthly frequency from 2002 to 2022. As shown in the top panel of Figure A1, the corn prices fluctuated between \$0.07/kg and \$0.30/kg, with an average of $0.15/kg.

Nitrogen prices were collected from two sources. The USDA National Agricultural Statistics Service provided yearly nitrogen fertilizer prices for the period 2002 to 2014. The DTN Retail Fertilizer Trends [@DTN2022] provided weekly nitrogen fertilizer prices from 2015 to 2022. To align with the monthly corn price data, the nitrogen price data from both sources were combined and converted into monthly data. Three most commonly used nitrogen fertilizer types were included: Liquid urea-ammonium nitrate with 28 percent nitrogen (UAN28), Liquid urea-ammonium nitrate with 32 percent nitrogen (UAN32), and anhydrous ammonia (ANHYD). All prices were all converted to the per kilogram (kg) of nitrogen basis. As shown in the middle panel of Figure A1, the nitrogen prices ranged from \$0.34/kg to \$2.37/kg, with an average of \$0.98/kg.

The nitrogen-corn price ratios were calculated as the nitrogen fertilizer price ($w$) divided by corn price ($p$), and the results were shown in the bottom panel of Figure A1. This study took the values at the 25th, 50th, and 75th percentiles of the historical monthly price ratios (pooled over the three types of nitrogen fertilizer) from 2002 to 2022. These three values were used to represent the low (5.44), middle (6.56), and high (7.67) price ratio scenarios. For consistency purposes, in the profitability calculations the corn price was selected at approximately the Spring 2022 price of \$0.28/kg. Consequently, the corresponding nitrogen prices for the three price ratio scenarios were approximately 5.44×0.28 = \$1.52/kg, 6.56×0.28 = \$1.84/kg, and 10.35×0.28 = \$2.15/kg. It is important to note that when different corn prices are used, the resulting profits of GWR will vary accordingly, but the variation will be proportional to the result of selected corn price and does not qualitatively alter the final conclusions.

# Appendix II: Results with Alternative Kernel Functions {-}

There are many different kernel functions to define the distance-decay weightings for GWR model. In the main text, the Gaussian kernel function was use, which can be mathematically represented as follows:

\begin{equation}
w_{ab}=exp⁡(-0.5[\frac{d_{ab}}{h}]^2 ),
(\#eq:q)
\end{equation}

where $d_{ab}$ denotes the Euclidean distance between observation $b$ and regression location $a$, $w_{ab}$ denotes the kernel weight assigned to observation $b$, and $h$ denotes the bandwidth.

This appendix compares several other alternative kernel functions, including the Exponential, Boxcar, Bi-square, and Tri-cube, in additional to the Gaussian. The functional forms and graphical representations of these kernels can be found in Figure A\@ref(fig:kernel). The estimation results obtained using alternative kernel functions, apart from the Gaussian kernel, are presented in Figures A\@ref(fig:value-kernel-bs) through A\@ref(fig:bias-kernel-boxcar).

```{r, fig.id = "kernel", fig.cap = "List of Kernel functions for GWR modeling. In the graphs of kernel function curves, the y-axis is the kernel weight, and the x-axis is the ratio of ($d_{ab}/h$). "}
knitr::include_graphics("kernel_function_list.png")
```

For the case of Exponential kernel, which is a continuous kernel like the Gaussian, the optimal bandwidth selection follows a similar corrected AIC approach and results in a similar bandwidth of 18 meters. For the discontinuous kernels of Boxcar, Bi-square, and Tri-cube, however, the 18-meter bandwidth does not encompass a sufficient number of neighboring observations for reliable local regressions, resulting in unrealistic estimation outcomes. To address this issue, an empirical bandwidth of 100 meters is chosen for these discontinuous kernels to enable the inclusion of an adequate number of local observations. Other bandwidth values (such as 80 meters, 150 meters) are also conducted and the resulting estimation outcomes are found to be comparable.

Overall, the economic evaluation results of GWR models presented in this study are not significantly influenced by the selection of the kernel function in GWR estimation. Despite slight variations in calculated profit and bias among the different kernels, the qualitative consistency of the final conclusions remains intact regardless of the chosen kernel.

## Comparing the values of GWR-based VRA and SCAM-based URA {-}

Figures A\@ref(fig:value-kernel-exp) through A\@ref(fig:value-kernel-tricube) show the distribution of the true economic value of GWR using four different kernels: Exponential, Boxcar, Bi-square, and Tri-cube.

\begin{equation}
\Delta^{GWR}=\Pi^{GWR}-\Pi^{SCAM}
(\#eq:label)
\end{equation}

Using the Gaussian kernel, we found that GWRR performs better as the price ratio increases, and GWRT performs better as the price ratio decreases (see Figure 3 of the main text). This holds true across all the other kernel function choices. Although there are some variations in the magnitudes of $\Delta^{GWR}$ among different kernel function choices, these differences do not significantly affect the conclusions drawn in the main text.

```{r, fig.id = "value-kernel-exp", fig.cap = "The true economic value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Exponential kernel function", fig.height = 3.5}
gg_appendix$g_value[[2]]
```

```{r, fig.id = "value-kernel-boxcar", fig.cap = "The true economic value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Boxcar kernel function", fig.height = 3.5}
gg_appendix$g_value[[3]]
```

```{r, fig.id = "value-kernel-bs", fig.cap = "The true economic value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Bi-square kernel function", fig.height = 3.5}
gg_appendix$g_value[[4]]
```

```{r, fig.id = "value-kernel-tricube", fig.cap = "The true economic value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Tri-cube kernel function", fig.height = 3.5}
gg_appendix$g_value[[5]]
```

## Bias in the estimation of the value of GWR-based VRA {-}

Figures A\@ref(fig:bias-kernel-exp) through A\@ref(fig:bias-kernel-tricube) show the distribution of the bias in the estimated economic value of GWR. Similar to the findings from using the Gaussian kernel presented in Figure 6 under the Gaussian kernel, the GWRR consistently exhibits significant over-estimation of the value of GWR-based VRA across all other kernel function choices. Also, the GWRT demonstrates smaller over-estimation at low price ratios compared to GWRR, while the over-estimation increases as price ratio rises. The presence of a similar fat tail in the GWRT bias distribution is also evident. The magnitudes of the estimation bias vary across kernel function choices, but the differences are not substantial and do not affect qualitatively our conclusions drawn in the main text.

```{r, fig.id = "bias-kernel-exp", fig.cap = "Bias in the estimation of the value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Exponential kernel function", fig.height = 3.5}
gg_appendix$g_bias[[2]]
```

```{r, fig.id = "bias-kernel-boxcar", fig.cap = "Bias in the estimation of the value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Boxcar kernel function", fig.height = 3.5}
gg_appendix$g_value[[3]]
```

```{r, fig.id = "bias-kernel-bs", fig.cap = "Bias in the estimation of the value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Bi-square kernel function", fig.height = 3.5}
gg_appendix$g_value[[4]]
```

```{r, fig.id = "bias-kernel-tricube", fig.cap = "Bias in the estimation of the value of GWR-based VRA for two GWR models (GWRR and GWRT) and under three nitrogen-corn price ratios (5.44 (low), 6.56 (middle), and 7.67 (high) $ kg^(-1)  / $ kg^(-1)), using Tri-cube kernel function", fig.height = 3.5}
gg_appendix$g_value[[5]]
```