### 2023-06-21 Codes/02_data_preparation.rmd

# Objectives

+ Create simulated on-farm experiments, and estimates site-specific
+ Generate yield response coefficients by GWR model
+ Price history data

# Preparation

```{r}
# === Packages ===#
library(sp)
library(spdep)
library(spatialreg)
library(sf)
library(raster)
library(data.table)
library(tidyverse)
library(dplyr)
library(magrittr)
library(gstat)
library(GWmodel)
library(scam)
library(mgcv)
library(magic)
library(stringr)
library(ggplot2)
library(tictoc)
library(here)
options(stringsAsFactors = FALSE)
```

Source all the functions in the Functions folder

```{r}
#--- load all the functions ---#
fs::dir_ls(here("GitControlled/Codes/Functions"), full.names = TRUE) %>%
  lapply(., function(x) source(x))

#--- load price history data ---#
price_ratio_data <- readRDS(here("Shared/Data/price_ratio_data.rds"))
```

# Generate data for simulation

## Create a field

```{r}
field_data <-
  # ! This is where you set the experiment parameters
  CJ(
    plot_length = 12, # the length of a plot (in number of cells)
    plot_width = 3, # the width of a plot (in number of cells)
    cell_buffer = 1,
    aunit_length = 2, # the length of an analysis unit (in number of cells)
    aunit_width = 3, # the width of an analysis unit (in number of cells)
    cell = 6, # the length of a cell in meter
    #* how wide the field is
    field_col = c(144), # the number of cell columns
    #* how tall the field is
    field_row = 72, # the number of row columns
    sp_range = c(600),
    gstat_model = "Sph"
  ) %>%
  rowwise() %>%
  mutate(
    field_sf = list(
      make_field(
        field_col = field_col,
        field_row = field_row,
        aunit_length = aunit_length,
        aunit_width = aunit_width,
        cell = cell,
        cell_buffer = cell_buffer
      )
    ),
    #* price ratio scenarios
    pCorn = 7 / 25.4, # $/kg
    pRatio_ls = list(
      price_ratio_data$pRatio %>%
        quantile(na.rm = TRUE, prob = c(0.05, 0.25, 0.5, 0.75, 0.95)) %>%
        round(digits = 2)
    )
  )

#* save the fields
saveRDS(field_data, here("Shared/Data/field_data.rds"))
```

## Add trial design layout

```{r}
field_data <- readRDS(here("Shared/Data/field_data.rds"))

#* (use the fixed Latin Square design)
field_with_design <-
  field_data %>%
  rowwise() %>%
  mutate(design_layout = list(
    make_design_layout(plot_length, field_col)
  )) %>%
  dplyr::select(-plot_length) %>%
  unnest(cols = "design_layout") %>%
  filter(design_name == "Latin Square Fixed") %>%
  rowwise() %>%
  mutate(plot_block_id_data = list(
    gen_plot_block_ids(
      field_sf = field_sf,
      plot_length = plot_length,
      plot_width = plot_width,
      cols_plot_in_block = cols_plot_in_block,
      rows_plot_in_block = rows_plot_in_block,
      cell_buffer = cell_buffer
    )
  )) %>%
  mutate(field_sf = list(
    left_join(field_sf, plot_block_id_data, by = "cell_id")
  )) %>%
  dplyr::select(-plot_block_id_data) %>%
  mutate(
    data_file_name =
      paste0(
        stringr::str_replace_all(design_name, " ", ""),
        "_", field_col, ".rds"
      ) %>%
        here("Shared", "Data", .)
  )

#* save the fields with designs
saveRDS(field_with_design, here("Shared/Data/field_with_design.rds"))
```

## Generate true field parameters

```{r}
field_with_design <- readRDS(here("Shared/Data/field_with_design.rds"))

field_parameters <-
  readRDS(here("Shared/Data/field_data.rds")) %>%
  mutate(field_pars = list(
    gen_field_pars(
      sp_range = sp_range,
      gstat_model = gstat_model,
      field_sf = field_sf,
      nsim = 1000
    )
  ))

#* save the field parameters
saveRDS(field_parameters, here("Shared/Data/field_parameters.rds"))
```

## Create regression data

```{r}

field_parameters <- readRDS(here("Shared/Data/field_parameters.rds"))

tic()
lapply(
  1:nrow(field_with_design),
  function(x) gen_reg_data(x, field_with_design, field_parameters)
)
toc()
```