### 2023-07-28 Codes/3_make_figures_publication.rmd
fix the issues of inconsistent description and figure on single cell bias

# Set figure theme 
```{r}
theme_fig <-
  theme_bw() +
  theme(
    axis.title.x =
      element_text(
        size = 11, angle = 0, hjust = .5, vjust = -0.3, family = "Times"
      ),
    axis.title.y =
      element_text(
        size = 11, angle = 90, hjust = .5, vjust = .9, family = "Times"
      ),
    axis.text.x =
      element_text(
        size = 8, angle = 0, hjust = .5, vjust = 1.5, family = "Times"
      ),
    axis.text.y =
      element_text(
        size = 8, angle = 0, hjust = 1, vjust = 0, family = "Times"
      ),
    axis.ticks =
      element_line(
        linewidth = 0.3, linetype = "solid"
      ),
    axis.ticks.length = unit(.15, "cm"),
    #--- legend ---#
    legend.text =
      element_text(
        size = 11, angle = 0, hjust = 0, vjust = 0.5, family = "Times"
      ),
    legend.title =
      element_text(
        size = 11, angle = 0, hjust = 0, vjust = 0, family = "Times"
      ),
    legend.key.size = unit(0.5, "cm"),
    #--- strip (for faceting) ---#
    strip.text = element_text(size = 8, family = "Times"),
    #--- plot title ---#
    plot.title = element_text(family = "Times", face = "bold", size = 11),
    #--- margin ---#
    # plot.margin = margin(0, 0, 0, 0, "cm"),
    #--- panel ---#
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA)
  )
```

# Visualization the field and spatial units

```{r}
# fig.id = "field-layout",
# fig.cap = "Simulated field layout with spatial unit definitions",
# fig.dim = c(6, 6)

field_with_design <- readRDS(here("Shared/Data/field_with_design.rds"))
field_sf <- field_with_design[1, ]$field_sf[[1]]

plot_sf <-
  field_sf %>%
  nest_by(plot_id) %>%
  mutate(data = list(
    st_union(data)
  )) %>%
  unnest() %>%
  st_as_sf()

block_sf <-
  field_sf %>%
  nest_by(block_id) %>%
  mutate(data = list(
    st_union(data)
  )) %>%
  unnest() %>%
  st_as_sf()

block_text_sf <- st_centroid(block_sf)

#* Field: plots and blocks
g_field <-
  ggplot() +
  geom_sf(data = plot_sf, size = 0.2, fill = NA) +
  geom_sf(data = dplyr::filter(plot_sf, plot_id == 264), fill = "red", alpha = 0.5) +
  geom_sf(data = block_sf, fill = NA, linewidth = 1.5) +
  annotate("text", x = 830, y = 74, label = "plot", color = "red", size = 4, family = "Times") +
  geom_segment(
    aes(x = 830, xend = 830, y = 30, yend = -80),
    arrow = arrow(length = unit(0.5, "cm")),
    linewidth = 1,
    color = "red"
  ) +
  geom_sf_text(
    data = block_text_sf,
    aes(label = paste0("block ", block_id)),
    size = 5,
    family = "Times"
  ) +
  theme_void() +
  theme(plot.title = element_text(family = "Times")) +
  ggtitle("Panel (a): Plots and blocks in an experimental field")

## inside a plot
plot_sf_focus <- dplyr::filter(field_sf, plot_id == 1)

subplot_sf <-
  plot_sf_focus %>%
  nest_by(aunit_id, buffer) %>%
  mutate(data = list(
    st_union(data)
  )) %>%
  unnest() %>%
  st_as_sf() %>%
  mutate(label = ifelse(
    buffer == 1,
    "buffer",
    paste0("subplot-", aunit_id - 1)
  )) %>%
  mutate(buf_or_not = ifelse(label == "buffer", "buffer", "subplot"))

subplot_text_sf <- st_centroid(subplot_sf)

site_sf <- plot_sf_focus[7, ]

g_inside_plot <-
  ggplot() +
  geom_sf(data = plot_sf_focus, size = 0.2, fill = NA) +
  geom_sf_text(
    data = subplot_text_sf,
    aes(label = label),
    size = 3,
    family = "Times"
  ) +
  geom_sf(data = subplot_sf, linewidth = 1.2, fill = NA) +
  geom_sf(data = site_sf, fill = "red", alpha = 0.3) +
  annotate("text", x = 39, y = 429, label = "cell", family = "Times") +
  scale_fill_discrete(name = "") +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    plot.title = element_text(family = "Times")
  ) +
  ggtitle("Panel (b): Subplots, buffers, and sites in a single plot")

g_layout <- g_field / g_inside_plot

ggsave(
  "GitControlled/Writing/Figures/g_layout.png",
  g_layout,
  width = 6,
  dpi = 600
)
```

# Experimental Design

```{r}

# fig.id = "field-N-design",
# fig.cap = "Experiment design of nitrogen (N) rates"

reg_data <-
  readRDS(here("Shared/Data/LatinSquareFixed_144.rds")) %>%
  pull(reg_data) %>%
  .[[1]] %>%
  .[1]

data <- reg_data$data[[1]]

N_levels <- reg_data$N_levels[[1]]

data <-
  data %>%
  .[, Nid := as.numeric(as.factor(Nid))] %>%
  .[, Ntg := N_levels[Nid]]

f <-
  left_join(field_sf, data[, .(aunit_id, Ntg)], by = "aunit_id") %>%
  data.table() %>%
  .[, Ntg := factor(Ntg)] %>%
  .[buffer == 1, Ntg := "buffer"] %>%
  .[, Ntg := factor(Ntg, levels = c("buffer", as.character(levels(Ntg)[-7])))] %>%
  st_as_sf()

g_exp <-
  ggplot() +
  geom_sf(data = f, aes(fill = (Ntg)), size = 0.1) +
  scale_fill_viridis_d(name = "N rate (kg/ha)", direction = -1) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, family = "Times"),
    legend.title = element_text(hjust = 0.5, family = "Times"),
    legend.position = "bottom",
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 10, family = "Times")
  )

ggsave(
  "GitControlled/Writing/Figures/g_exp.png",
  g_exp,
  width = 6,
  dpi = 600
)
```

# Read the results

```{r}
results <-
  here("Shared/Results/gaussian/pi_data.rds") %>%
  readRDS() %>%
  .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
  .[, bias := pi_diff_est - pi_diff] %>%
  # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
  .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
  # === label price ratio
  .[, pLabelName := "Price Ratio (N/corn)"] %>%
  .[pRatio == 5.44, pLabel := "Low (5.44)"] %>%
  .[pRatio == 6.56, pLabel := "Middle (6.56)"] %>%
  .[pRatio == 7.67, pLabel := "High (7.67)"] %>%
  .[, pLabel := factor(pLabel, levels = c("Low (5.44)", "Middle (6.56)", "High (7.67)"))]
```

# The value of GWR-based VRA over SCAM-based URA for GWR-R and GWR-T

```{r}
# fig.id = "pi-dif-dist",
# fig.cap = "The value of GWR-based VRA over SCAM-based URA for GWR-R and GWR-T"
mean_data_value <-
  results %>%
  .[, .(pi_diff = median(pi_diff)),
    by = c("field_col", "pLabel", "type")
  ]

# saveRDS(mean_data_value, "Shared/Results/mean_data_value.rds")

g_value <-
  results %>%
  ggplot(data = .) +
  geom_histogram(
    aes(x = pi_diff),
    fill = NA,
    color = "blue",
    bins = 50,
    linewidth = 0.2
  ) +
  geom_vline(
    data = mean_data_value,
    aes(xintercept = pi_diff),
    color = "red",
    linewidth = 0.5
  ) +
  geom_text(
    data = mean_data_value, color = "red",
    aes(
      x = ifelse(pi_diff < 65, pi_diff + 5, pi_diff - 20), y = 150,
      label = paste0("Median = ", round(pi_diff, 2))
    ),
    angle = 0,
    hjust = -0.1,
    vjust = 0,
    size = 3,
    family = "Times"
  ) +
  facet_grid(pLabel ~ type) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks = 20 * (-2:3), limits = c(-45, 60)) +
  xlab("The value of VRA over URA ($ per ha)") +
  ylab("Number of Simulation Cases") +
  theme_fig

ggsave(
  "GitControlled/Writing/Figures/g_value.png",
  g_value,
  width = 6,
  height = 5,
  dpi = 600
)
```

# EONR estimation tendency

```{r}
# ggplot(est_data) +
# geom_histogram(aes(x = opt_N_scam)) +
# facet_grid(round(pN, digits = 2) ~ .)

est_data <-
  readRDS(here("Shared/Results/Gaussian/est_data.rds")) %>%
  unnest() %>%
  data.table() %>%
  .[, type := ifelse(transfer == 0, "GWRR", "GWRT")]

eorn_ratio_data <-
  est_data %>%
  .[, pRatio := pN / pCorn] %>%
  # only display the 25% (5.44), 50% (6.56), 75% (7.67) price ratio
  .[pRatio %in% c(5.44, 6.56, 7.67), ] %>%
  # === label price ratio
  .[, pLabelName := "Price Ratio (N/corn)"] %>%
  .[pRatio == 5.44, pLabel := "Low (5.44)"] %>%
  .[pRatio == 6.56, pLabel := "Middle (6.56)"] %>%
  .[pRatio == 7.67, pLabel := "High (7.67)"] %>%
  .[, pLabel := factor(pLabel, levels = c("Low (5.44)", "Middle (6.56)", "High (7.67)"))] %>%
  .[, .(eonr_ratio = mean(opt_N_gwr / opt_N_scam)), by = .(type, sim, pLabel)]

g_eonr_bias <-
  ggplot(data = eorn_ratio_data) +
  geom_histogram(
    aes(x = eonr_ratio),
    color = "blue",
    fill = NA,
    bins = 50,
    size = 0.2
  ) +
  geom_vline(xintercept = 1, color = "red") +
  facet_grid(pLabel ~ type) +
  ylab("Number of Simulation Cases") +
  xlab("Average Ratio of Estimated EONR to true EONR") +
  theme_fig

ggsave(
  "GitControlled/Writing/Figures/g_eonr_bias.png",
  g_eonr_bias,
  width = 6,
  height = 5,
  dpi = 600
)
```

# Comparison of Estimated and True Coefficients

```{r}

# fig.id = "true-vs-estimated-coef-gwr-r",
# fig.cap = "Comparison of Estimated and True Coefficients"

#* read the simulation data for a sigle simulationused for illustration
single_sim <-
  here("Shared/Results/Gaussian/aunit_sim_single.rds") %>%
  readRDS() %>%
  .[, type := ifelse(transfer == 0, "GWRR", "GWRT")]

plot_data <-
  single_sim %>%
  .[type == "GWRR", ] %>%
  .[, .(aunit_id, b1, b2, b1_hat, b2_hat)] %>%
  melt(id.var = "aunit_id") %>%
  .[, type := ifelse(str_detect(variable, "hat"), "Estimated", "True")] %>%
  .[, variable := gsub("_hat", "", variable)] %>%
  dcast(aunit_id + variable ~ type, value.var = "value")

g_b1 <-
  plot_data[variable == "b1", ] %>%
  ggplot(data = .) +
  geom_point(aes(y = Estimated, x = True), size = 0.3) +
  xlim(0, NA) +
  ylim(0, NA)

g_b2 <-
  plot_data[variable == "b2", ] %>%
  ggplot(data = .) +
  geom_point(aes(y = Estimated, x = True), size = 0.3) +
  geom_hline(yintercept = 0, color = "red") +
  xlim(NA, 0.1) +
  ylim(NA, 0.1) +
  theme_fig

g_comp_coef <- g_b1 / g_b2
```

# Comparison of Estimated and True EONR

```{r}
# fig.id = "true-vs-estimated-optn-gwr-r",
# fig.cap = "Comparison of Estimated and True EONR"

# est_data[sim == 1 & pn_pc == 10.35, ] %>%
#   .[, type := ifelse(transfer == 0, "GWRR", "GWRT")] %>%
#   ggplot(data = .) +
#     geom_histogram(aes(x = opt_N_gwr)) +
#     facet_grid(. ~ type)

g_comp_eonr <-
  single_sim %>%
  ggplot(data = .) +
  geom_point(aes(y = opt_N_gwr, x = opt_N), size = 0.3) +
  geom_abline(slope = 1, color = "red") +
  xlab("True Optimal Nitrogen Rate (kg/ha)") +
  ylab("Estimated Optimal Nitrogen Rate (kg/ha)") +
  facet_grid(. ~ type) +
  theme_fig +
  coord_equal()

ggsave(
  "GitControlled/Writing/Figures/g_comp_eonr.png",
  g_comp_eonr,
  width = 6,
  height = 5,
  dpi = 600
)
```

# Bias in the estimation of the value of GWR-based VRA over SCAM-based URA

```{r}
# fig.id = "bias-est-pi",
# fig.cap = "Bias in the estimation of the value of GWR-based VRA over SCAM-based URA for GWR-R and GWR-T"

# mean_data_bias[type == "GWRT" & pLabel == 10.35, bias] %>% hist

median_bias_data <-
  results %>%
  .[, bias := pi_diff_est - pi_diff] %>%
  .[, .(bias = median(bias)),
    by = c("field_col", "pLabel", "type")
  ] %>%
  print()

g_bias <-
  results %>%
  .[, bias := pi_diff_est - pi_diff] %>%
  .[bias < 150, ] %>%
  ggplot(data = .) +
  geom_histogram(
    aes(x = bias),
    fill = NA,
    color = "blue",
    size = 0.2,
    bins = 50
  ) +
  geom_vline(
    data = median_bias_data, aes(xintercept = bias),
    color = "red",
    size = 0.5
  ) +
  geom_text(
    data = median_bias_data, color = "red",
    aes(
      x = ifelse(bias < 65, bias + 2, bias - 85), y = 100,
      label = paste0("Median = ", round(bias, 2)),
      family = "Times"
    ),
    angle = 0, hjust = -0.1, vjust = 0, size = 3
  ) +
  facet_grid(pLabel ~ type) +
  scale_y_continuous(expand = c(0, 0)) +
  xlab("Bias in the Estimation of the Value of VRA over URA ($ per ha)") +
  ylab("Number of Simulation Cases") +
  theme_fig

ggsave(
  "GitControlled/Writing/Figures/g_bias.png",
  g_bias,
  width = 6,
  height = 5,
  dpi = 600
)
```

# The cause of significant over-estimation of the value of GWR-based VRA

```{r}

# fig.id = "why-bias-many",
# fig.cap = "The cause of significant over-estimation of the value of GWR-based VRA"

# /*+++++++++++++++++++++++++++++++++++
#' # Prepare parameters
# /*+++++++++++++++++++++++++++++++++++
field_parameters <- readRDS(here("Shared/Data/field_parameters.rds"))

pCorn <- field_parameters$pCorn
pN <- pCorn * field_parameters$pRatio_ls[[1]][2] # medium price scenario

# /*+++++++++++++++++++++++++++++++++++
#' # Prepare yield response curves
# /*+++++++++++++++++++++++++++++++++++
il_data_oe <-
  readRDS(here("Shared/Results/Gaussian/il_data_oe.rds")) %>%
  .[,
    lapply(.SD, mean),
    by = aunit_id,
    .SDcols = c(
      "b0", "b1", "b2", "Nk",
      "b0_hat", "b1_hat", "b2_hat",
      "opt_N", "opt_N_gwr", "opt_N_scam"
    )
  ]

n_data <-
  data.table(
    N = seq(
      min(il_data_oe$opt_N_gwr) - 50,
      max(il_data_oe$opt_N_gwr, il_data_oe$opt_N_scam),
      length = 30
    )
  )

gwr_curv_data <-
  expand_grid_df(il_data_oe, n_data) %>%
  .[, yield := (b0_hat + b1_hat * N + b2_hat * N^2) / 1000] %>%
  .[, profit := pCorn * yield * 1000 - pN * N] %>%
  .[, y_hat_gwr := (b0_hat + b1_hat * opt_N_gwr + b2_hat * opt_N_gwr^2) / 1000] %>%
  .[, y_hat_scam := (b0_hat + b1_hat * opt_N_scam + b2_hat * opt_N_scam^2) / 1000]

single_aunit <- 
  gwr_curv_data[abs(opt_N-150) < 10, ] %>%
  .[which.max(yield), aunit_id]

set.seed(710527)
aunit_id_ls <- c(sample(il_data_oe$aunit_id, 49), single_aunit)

gwr_curv_data <- gwr_curv_data[aunit_id %in% aunit_id_ls, ]

# /*+++++++++++++++++++++++++++++++++++
#' # Prepare point data
# /*+++++++++++++++++++++++++++++++++++

point_data <-
  gwr_curv_data %>%
  .[aunit_id %in% aunit_id_ls, ] %>%
  .[, .(aunit_id, y_hat_gwr, y_hat_scam, opt_N_gwr, opt_N_scam)] %>%
  unique(by = "aunit_id") %>%
  melt(id.var = "aunit_id") %>%
  .[, var_type := fifelse(str_detect(variable, "y_hat"), "yield", "N")] %>%
  .[, type := fifelse(str_detect(variable, "gwr"), "GWR", "SCAM")] %>%
  .[, variable := NULL] %>%
  data.table::dcast(aunit_id + type ~ var_type, value.var = "value") %>%
  .[, profit := pCorn * yield * 1000 - pN * N]

# /*+++++++++++++++++++++++++++++++++++
#' # Figure
# /*+++++++++++++++++++++++++++++++++++
g_why_bias_many <-
  ggplot() +
  geom_line(
    data = gwr_curv_data,
    aes(y = yield, x = N, group = aunit_id),
    size = 0.3,
    color = "grey"
  ) +
  geom_point(
    data = point_data,
    aes(y = yield, x = N, color = type),
    size = 0.6
  ) +
  ylab("Estimated Yield (ton/ha)") +
  xlab("Nitrogen Rate (kg/ha)") +
  theme_fig +
  theme(
    legend.position = "bottom"
  ) +
  scale_color_discrete(name = "Estimated EONR")

ggsave(
  "GitControlled/Writing/Figures/g_why_bias_many.png",
  g_why_bias_many,
  width = 6,
  height = 5,
  dpi = 600
)
```

# An illustration of over-estimation of the value of GWR-based VRA over SCAM-based URA

```{r}
# fig.id = "why-bias-single",
# fig.cap = "An illustration of over-estimation of the value of GWR-based VRA over SCAM-based URA",
# fig.dim = c(6, 7)

# /*+++++++++++++++++++++++++++++++++++
#' # Preapare yeld response curves
# /*+++++++++++++++++++++++++++++++++++

true_curv_data <-
  il_data_oe %>%
  .[aunit_id %in% single_aunit, ] %>%
  expand_grid_df(., n_data) %>%
  .[, y_true_curve_raw := (b0 + b1 * N + b2 * N^2) / 1000] %>%
  .[, y_true_curve_pl := (b0 + b1 * Nk + b2 * Nk^2) / 1000] %>%
  .[, yield := fifelse(N < Nk, y_true_curve_raw, y_true_curve_pl)] %>%
  .[, profit := pCorn * yield * 1000 - pN * N] %>%
  .[, .(yield, profit, N)] %>%
  .[, type := "True"]

gwr_curv_data_f <-
  gwr_curv_data[aunit_id %in% single_aunit, ] %>%
  .[, .(yield, profit, N)] %>%
  .[, type := "GWR-estimated"]

curv_data <-
  rbind(true_curv_data, gwr_curv_data_f) %>%
  .[, type := factor(type, levels = c("True", "GWR-estimated"))]

# /*+++++++++++++++++++++++++++++++++++
#' # Prepare yield points data
# /*+++++++++++++++++++++++++++++++++++
true_yield_data <-
  il_data_oe %>%
  .[aunit_id %in% single_aunit, ] %>%
  .[, .(b0, b1, b2, Nk, opt_N_gwr, opt_N_scam, opt_N)] %>%
  melt(id.vars = c("b0", "b1", "b2", "Nk")) %>%
  setnames("value", "N") %>%
  .[, y_true_raw := (b0 + b1 * N + b2 * N^2) / 1000] %>%
  .[, y_true_pl := (b0 + b1 * Nk + b2 * Nk^2) / 1000] %>%
  .[, yield := fifelse(N < Nk, y_true_raw, y_true_pl)] %>%
  .[, profit := pCorn * yield * 1000 - pN * N] %>%
  .[, type := fcase(
    variable == "opt_N_gwr", "GWR",
    variable == "opt_N_scam", "SCAM",
    variable == "opt_N", "True"
  )] %>%
  .[, .(yield, profit, N, type)]

point_data_gwr <-
  point_data[aunit_id %in% single_aunit, ]

# /*+++++++++++++++++++++++++++++++++++
#' # Figure
# /*+++++++++++++++++++++++++++++++++++
g_yield <-
  ggplot() +
  geom_line(data = curv_data, aes(y = yield, x = N, linetype = type)) +
  geom_point(
    data = true_yield_data,
    aes(y = yield, x = N, color = type, shape = type)
  ) +
  geom_point(
    data = point_data_gwr,
    aes(y = yield, x = N, color = type, shape = type)
  ) +
  ylab("Yield (ton/ha)") +
  xlab("Nitrogen Rate (kg/ha)") +
  scale_color_discrete(name = "EONR") +
  scale_shape_discrete(name = "EONR") +
  scale_linetype_discrete(name = "Yield Response Fnctions") +
  ggtitle("(a) Estimated and True Yields") +
  theme_fig

g_profit <-
  ggplot() +
  geom_line(data = curv_data, aes(y = profit / 1000, x = N, linetype = type)) +
  geom_point(
    data = true_yield_data,
    aes(y = profit / 1000, x = N, color = type, shape = type)
  ) +
  geom_point(
    data = point_data_gwr,
    aes(y = profit / 1000, x = N, color = type, shape = type)
  ) +
  ylab("Partial Profit ($1000/ha)") +
  xlab("Nitrogen Rate (kg/ha)") +
  scale_color_discrete(name = "EONR") +
  scale_shape_discrete(name = "EONR") +
  scale_linetype_discrete(name = "Profit Response Fnctions") +
  ggtitle("(b) Estimated and True Profits") +
  theme_fig

g_why_bias_single <- g_yield / g_profit

ggsave(
  "GitControlled/Writing/Figures/g_why_bias_single.png",
  g_why_bias_single,
  width = 6.5,
  dpi = 600
)
```