### 2023-11-05 Writing/JAAEA/2ndRoundRevision/reply_to_reviewer_2.rmd
replies started

---
title: "Reply"
output: 
  html_document: default
---

Your comments are in blue.  Our responses are in black.

# General Comments

+ <span style = "color: blue;">I still believe this research is related to studies that examine the economic value of forecasting models by assessing the economic effects stemming from inaccuracy (e.g.,
reduction in a business’ profits by assuming a mean-squared-error loss function instead of an
asymmetric loss, for example). Although there have been improvements in references to the
agronomy and statistical literatures, it would still be useful to add 1-2 sentences in the
conclusion that reference the broader economics literature on this topic. </span>

Response: We do not address this as you suggested.

+ <span style = "color: blue;"> As mentioned in my previous review, the term "economically optimal" has different meanings among agricultural economists. The new footnote (footnote 1 on p. 5) is appreciated, but here you should specify that "economically optimal" in the context of the article does not mean "socially optimal."</span>

Response:

+ <span style = "color: blue;"> None of the figure numbers appear in the text; instead, each figure reference is followed by "Error! Reference source not found." These all should be corrected. </span>

Response: We apologize for these mistakes. We have corrected theses errors.

+ In Section 3, to reduce confusion, you should consider using the word "Simulated" in the
section heads (i.e., 3.1 Simulated Field Layout, 3.2 Simulated True Cell-Specific Yield
Response Function, 3.3 Simulated Trial Design, and 3.4 Simulated Yield Data).

Response:

+ All of the figures have their dimensions and DPI listed (1422x812mm). This information is
not necessary and should have been removed prior to submission.

Response:

Specific Comments

+ P. 4, lines 58-59: This statement, while perhaps likely, is just speculation. The authors should
either check with Trevisan et al. (2021) that this was the reason (and then remove "It seems
likely") or just delete the sentence entirely.
o The same comment applies to P. 8, lines 134-136.

Response:

+ P. 6, lines 91-95: Implicitly, you seem to be suggesting that the set of points from which you
would take (input rate, yield) data is smaller than n, the total number of sites associated with
the characteristics vector. Either way, this should be clarified in the text.

Response:

+ P. 9, lines 146-148: This seems overly negative of Trevisan et al. (2021), as well as clunky
wording ("provided an example at step 3"). Since you say this assumption is unavoidable, it is done by all studies, including the ones you mention above: Anselin et al, Bullock et al, Liu et al, and Swinton and Lowenberg-DeBoer. You should remove this sentence entirely because it really adds nothing. Later on in the paragraph, you address the thing that's more important: inaccurate estimation of yield response using GWR, because it exaggerates the true heterogeneity in the yield response functions, leads to comparatively worse bias.

Response:

+ P. 10, lines 185-186: This almost sounds misleading; "were available for analysis" seems to
suggest you were using real-world data. You should re-write this as "In total, simulated data from 1,440 subplots were used for analysis."

Response:

+ P. 10, line 199: In this sentence, you should mention which specific statistical distribution the
error term follows, despite the fact that you mention Gaussian random fields below.

Response:

+ P. 12, line 212: It would be useful to add a footnote here explaining why the range of 600m
was chosen.

Response:

+ P. 12, line 229: Given that JAAEA is an economics journal and not an agronomy journal, you
should rewrite "Latin square trial design schema" as "trial design" here, and at the end of the
next sentence, write "each column of the block (i.e., a Latin square)."

Response:

+ P. 14, line 254: For additional clarity, you could point out that the EOUNA rate is just the
value of N that solves 𝑝 ⋅ 𝑓(𝑁) - 𝑤 ⋅ 𝑁, where 𝑓(𝑁) is SCAM-estimated.

Response:

+ P. 15, lines 277-279: Add a sentence (perhaps in a footnote) mentioning how this shape
constraint is actually imposed.

Response:

+ PP. 22-23, lines 439-441: This sentence should be removed because it seems speculative.
More importantly, researchers do not "have" to implement a methodology because it is
conveniently available in standard software; instead, they could code their own models,
methods, etc.

Response:

+ P. 33, Figure 2: In the notes to this figure, you should add that the 1,440 subplots are
illustrated as 24 rows x 12 columns, each containing 5 subplots (24 x 12 x 5 = 1,440). Also
add that there are 72 rows x 120 columns of cells (72 x 120 = 8,640).

Response:

Writing

+ P. 4, line 45: Replace "dramatic" with "seemingly implausibly large"

Response:

+ P. 4, line 54: Replace "suspect" with "hypothesize"
o Same comment for "suspect" on P. 9, line 153

Response:

+ P. 5, line 77: You do not need to write (OFPE) again here since you have already introduced
the acronym.

Response:

+ P. 8, line 124: You should place the first instance of "(EONR)" here after "economically
optimal nitrogen rates" rather than line 132.

Response:

+ P. 8, line 130: Replace "dangerous" with "problematic"
o Same comment for P. 22, line 438

Response:

+ P. 8, line 145: Remove the word "the" in front of "step"

Response:

+ P. 11, lines 200-201: Replace "are desirable properties" with "is a desirable property"

Response:

+ P. 13, line 233: Replace "due the" with "due to the"

Response:

+ P. 14, line 262: This is the first and only mention in the paper of a benchmark. Given that,
what is a "lower" benchmark?

Response:

+ P. 16, line 295: Replace "as" with "the"

Response:

+ P. 19, line 269: Replace "truea" with "true"

Response:

+ P. 22, line 453: Remove the word "please"

Response:

+ P. 24, lines 468-474: Remove the clause "After the development of software that..."
Additionally, you should add quotation marks to both instances of "pure" in this paragraph.
Response: