### 2023-11-05 Writing/JAAEA/2ndRoundRevision/reply_to_reviewer_2.rmd
modified

---
title: "Reply"
output: 
  html_document: default
---

Your comments are in blue.  Our responses are in black.

# General Comments

+ <span style = "color: blue;">I still believe this research is related to studies that examine the economic value of forecasting models by assessing the economic effects stemming from inaccuracy (e.g.,
reduction in a business’ profits by assuming a mean-squared-error loss function instead of an
asymmetric loss, for example). Although there have been improvements in references to the
agronomy and statistical literatures, it would still be useful to add 1-2 sentences in the
conclusion that reference the broader economics literature on this topic. </span>

Response: We do not address this as you suggested.

+ <span style = "color: blue;"> As mentioned in my previous review, the term "economically optimal" has different meanings among agricultural economists. The new footnote (footnote 1 on p. 5) is appreciated, but here you should specify that "economically optimal" in the context of the article does not mean "socially optimal."</span>

Response:

+ <span style = "color: blue;"> None of the figure numbers appear in the text; instead, each figure reference is followed by "Error! Reference source not found." These all should be corrected. </span>

Response:

+ <span style = "color: blue;"> In Section 3, to reduce confusion, you should consider using the word "Simulated" in the section heads (i.e., 3.1 Simulated Field Layout, 3.2 Simulated True Cell-Specific Yield Response Function, 3.3 Simulated Trial Design, and 3.4 Simulated Yield Data). </span>

Response: Instead of making the suggested changes, we simply renamed the Methods section to "Methods: Monte Carlo Simulations," so that it is clear that all the subsequent sections are about simulations.

+ <span style = "color: blue;"> All of the figures have their dimensions and DPI listed (1422x812mm). This information is not necessary and should have been removed prior to submission. </span>

Response: We apologize this mistake. This has been fixed in this version of the manuscript.

# Specific Comments

+ <span style = "color: blue;"> P. 4, lines 58-59: This statement, while perhaps likely, is just speculation. The authors should either check with Trevisan et al. (2021) that this was the reason (and then remove "It seems likely") or just delete the sentence entirely. The same comment applies to P. 8, lines 134-136. </span>

Response: You are correct and we have removed the sentence entirely.

+ <span style = "color: blue;">P. 6, lines 91-95: Implicitly, you seem to be suggesting that the set of points from which you would take (input rate, yield) data is smaller than n, the total number of sites associated with the characteristics vector. Either way, this should be clarified in the text.</span>

Response: We have modified the sentence to the following so that it describes what GWR does more accurately:

  - Old: "Alternatively, a site-specific yield response function might be estimated simply by taking (input rate, yield) data from points around a site of interest"
  - New: "Alternatively, a site-specific yield response function might be estimated simply by giving greater weights to the points around a site of interest"

+ <span style = "color: blue;"> P. 9, lines 146-148: This seems overly negative of Trevisan et al. (2021), as well as clunky wording ("provided an example at step 3"). Since you say this assumption is unavoidable, it is done by all studies, including the ones you mention above: Anselin et al, Bullock et al, Liu et al, and Swinton and Lowenberg-DeBoer. You should remove this sentence entirely because it really adds nothing. Later on in the paragraph, you address the thing that's more important: inaccurate estimation of yield response using GWR, because it exaggerates the true heterogeneity in the yield response functions, leads to comparatively worse bias. </span>

Response: We dropped the sentence.

+ <span style = "color: blue;"> P. 10, lines 185-186: This almost sounds misleading; "were available for analysis" seems to suggest you were using real-world data. You should re-write this as "In total, simulated data from 1,440 subplots were used for analysis." </span>

Response: We edited the sentence according to your suggestion.

+ <span style = "color: blue;"> P. 10, line 199: In this sentence, you should mention which specific statistical distribution the error term follows, despite the fact that you mention Gaussian random fields below. </span>

Response: It follows a normal distribution in a way it is spatially correlated. This has been made clear in the manuscript (see page ??, line ??).

+ <span style = "color: blue;">P. 12, line 212: It would be useful to add a footnote here explaining why the range of 600m was chosen.</span>

Response: We added an explanation that the 600m range was chosen based on empirical data of our existing on-farm trials (see page ??, lines ??).

+ <span style = "color: blue;">P. 12, line 229: Given that JAAEA is an economics journal and not an agronomy journal, you should rewrite "Latin square trial design schema" as "trial design" here, and at the end of the next sentence, write "each column of the block (i.e., a Latin square)." </span>

Response: We followed your suggestions and texts were modified accordingly.

+ <span style = "color: blue;">P. 14, line 254: For additional clarity, you could point out that the EONR rate is just the value of N that solves $\\cdot f(N) - \omega\dcot N$, where $f(N)$ is SCAM-estimated. </span>

Response:

+ <span style = "color: blue;"> P. 15, lines 277-279: Add a sentence (perhaps in a footnote) mentioning how this shape constraint is actually imposed. </span>

Response: Expalining how the shape constraint is imposed would take more than a page of mathematical explanations. We added a footnote that  cites Pya's journal article titled "Shape constrained additive models" published in Statistics and computing.

+ <span style = "color: blue;">PP. 22-23, lines 439-441: This sentence should be removed because it seems speculative. More importantly, researchers do not "have" to implement a methodology because it is conveniently available in standard software; instead, they could code their own models, methods, etc.  </span>

Response: There is no speculation in this sentence. We are stating a drawback of linear-plateau model. What we state is a irrefutable fact and important in real world practice of estimating yield response function.

+ <span style = "color: blue;">P. 33, Figure 2: In the notes to this figure, you should add that the 1,440 subplots are illustrated as 24 rows x 12 columns, each containing 5 subplots (24 x 12 x 5 = 1,440). Also add that there are 72 rows x 120 columns of cells (72 x 120 = 8,640).  </span>

Response:

# Writing

+ <span style = "color: blue;">P. 4, line 45: Replace "dramatic" with "seemingly implausibly large"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 4, line 54: Replace "suspect" with "hypothesize". Same comment for "suspect" on P. 9, line 153</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 5, line 77: You do not need to write (OFPE) again here since you have already introduced
the acronym.</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 8, line 124: You should place the first instance of "(EONR)" here after "economically
optimal nitrogen rates" rather than line 132.</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 8, line 130: Replace "dangerous" with "problematic" Same comment for P. 22, line 438 </span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 8, line 145: Remove the word "the" in front of "step"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 11, lines 200-201: Replace "are desirable properties" with "is a desirable property" </span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 13, line 233: Replace "due the" with "due to the"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 14, line 262: This is the first and only mention in the paper of a benchmark. Given that, what is a "lower" benchmark?</span>

Response: We have changed the wording to "In this scenario."

+ <span style = "color: blue;">P. 16, line 295: Replace "as" with "the"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 19, line 269: Replace "truea" with "true"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 22, line 453: Remove the word "please"</span>

Response: We followed your suggestion.

+ <span style = "color: blue;">P. 24, lines 468-474: Remove the clause "After the development of software that..." Additionally, you should add quotation marks to both instances of "pure" in this paragraph.</span>

Response: We followed your suggestion.